---
title: "<center>Reproducible Area Deprivation Index (ReADI | 2015)</center>"
author: "<br><center>[Nicole Gladish, PhD](https://profiles.stanford.edu/nicole-gladish)</center>"
date: "<center>`r format(Sys.time(), '%d %B, %Y')`</center>"
output:
  html_document:
    includes:
      after_body: "../templates/NG_footer.html"
      in_header: "../templates/header.html"
    toc: true
    toc_depth: 2
    number_sections: true
    toc_float: 
      collapsed: true
      smooth_scroll: true
    css: "../templates/bootstrap.css"
    self_contained: true
---
<br><br>

```{css, echo = FALSE}
  /* General code block styling */
  pre {
    background-color: #2e2e2e;
    color: #f5f5f5;
    padding: 8px 12px;
    border-radius: 4px;
    font-family: "Courier New", monospace;
  }
  code {
      background-color: #2e2e2e;
    color: #f5f5f5;
    padding: 8px 12px;
    border-radius: 4px;
    font-family: "Courier New", monospace;
}
  .hljs{
  display: block;
  padding: 0.5em;
  background: #2e2e2e;
  color: #f5f5f5;
  }

  /* Strings */
  .hljs-string {
      color: #c2a3fd; 
  }

  /* Numbers */
  .hljs-number, 
  .hljs-literal,
  .hljs-logical,
  .hljs-built_in, 
  .hljs-variable.constant,
  .hljs-keyword, 
  .hljs-variable.language, 
  .hljs-title.function{
      color: #8fff84; 
  }

  /* Comments */
  .hljs-comment, .hljs-quote {
      color: #ff7777; 
  }

  /* Symbols and punctuation (e.g., parentheses, brackets) */
.hljs-punctuation, .hljs-operator {
      color: #13cee0; 
  }
  
  /* General card block styling */
  
div.card {
    background-color: #71797E; /* Dark background color */
    border-radius: 10px;
    padding: 10px;
    color: white; /* Set text color to white */
    margin-bottom: 0;
    margin-top: 10px; /* Add some space above paragraphs if needed */
  }
   /* Explicitly set all header text inside .card to white (in case bootstrap overrides it) */
  div.card h1, div.card h2, div.card h3, div.card h4, div.card h5, div.card h6 {
    color: white; /* Set header text color to white */
  }
```

```{r setup, include=FALSE}
## Warn if R version is older than what was used for this analysis
if (getRversion() < "4.5.0") {
  warning(
    "This analysis was developed under R 4.5.0 ",
    "You are using R ", as.character(getRversion()),
    ". If you see package errors, consider upgrading to at least R 4.5.0"
  )
}

## Check for required packages and fail with a clear message if missing
needed_pkgs <- c(
  "tidyverse",
  "censusapi",
  "reshape2",
  "plyr",
  "SciViews",
  "readxl",
  "tidycensus",
  "psych",
  "tigris",
  "spdep",
  "statar",
  "dbplyr"
)

missing_pkgs <- needed_pkgs[
  !vapply(needed_pkgs, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_pkgs) > 0) {
  msg <- paste0(
    "The following R packages are missing: ",
    paste(missing_pkgs, collapse = ", "), ".\n\n",
    "Please run renv::restore() in the project root for this year\n",
    "before knitting this document.\n\n",
    "If renv::restore() fails, see the README for this year\n",
    "for operating system–specific setup instructions."
  )
  stop(msg, call. = FALSE)
}

## Global options
options(tigris_use_cache = TRUE)
```

# Brief Overview of ADI and ReADI
***
<br>

The Area Deprivation Index (ADI) was originally developed by Singh (2003) as a factor-based measure of area-level socioeconomic deprivation using 17 census-derived indicators. Kind et al. (2014, 2018) later operationalized the ADI for broader use through the Neighborhood Atlas, providing block-group level scores for multiple ACS periods.

However, subsequent work has shown that the Neighborhood Atlas ADI (NA-ADI) implementation has several methodological limitations (e.g., reuse of 1990 factor weights on newer data, application of weights to unstandardized variables, and reliance on outdated variable thresholds). These issues affect reproducibility and interpretability.

The Reproducible Area Deprivation Index (ReADI) is a fully documented re-implementation of the ADI that:
- Standardizes all input variables and derives factor scores directly from the factor analysis output.
- Re-estimates factor weights for each ACS period using a one-factor model with population weights.
- Modernizes key variable definitions and thresholds (e.g., education and income disparity) to reflect current socioeconomic conditions.
- Incorporates population size in the factor model to account for highly skewed area population distributions.

For full methodological details, see the main ReADI methods documentation and website.

# ReADI Preparation
***

> Libraries

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code294">Show Code</button>
  <div id="code294" class="collapse">

```{r, results = F, message = F}
library(tidyverse)
library(censusapi)
library(reshape2) 
library(plyr) 
library(dplyr) 
library(SciViews) 
library(readxl) 
library(tidycensus) 
library(psych)
options(tigris_use_cache = TRUE)
library(tigris)
library(spdep)
library(statar)
```

```{r}
sessionInfo()
```
  </div>
</div>
  
<br><br>

## Retrieving 2011-2015 5-Year ACS
***

> Obtaining Census Key

In order to obtain the census API data you are required to obtain a census key from [here](https://api.census.gov/data/key_signup.html).

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code339">Show Code</button>
  <div id="code399" class="collapse">

```{r, eval = F}
if (identical(Sys.getenv("CENSUS_API_KEY") != "", F)) {
  stop(
    "Census API key not found.\n",
    "Please set the CENSUS_API_KEY environment variable before running this document.\n"
  )
}
states <- setdiff(unique(fips_codes$state_code), c("60", "66", "69", "74", "78")) # Removing US territories
```
  </div>
</div>
  
<br><br>

### Pulling Variables Required
***

There are some areas which have no population or are located in water etc. These should be removed so they are not influencing the imputation of other regions. More information [here](https://www2.census.gov/geo/pdfs/maps-data/data/tiger/tgrshp2020pl/TGRSHP2020PL_TechDoc_Ch4.pdf). We are pulling data from three geographies: county, census tract, and census block group. Definitions on what these are can be found in this [census documentation](https://www2.census.gov/geo/pdfs/reference/GARM/Ch10GARM.pdf).

Here is the breakdown of how many of each area exist:

<div class = "row">
<div class = "col-md-4">
|Area       |Amount   |
|:----------|:--------|
|CBG        |218,940  |
|CT         |73,297   |
|County     |3,142    |
</div>

<div class = "col-md-8">
</div>
</div>

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code357">Show Code</button>
  <div id="code357" class="collapse">
```{r, eval = F}
ADI_Vars <- c(
       "B01003_001E",
       "B19013_001E",
       "B19001_002E", "B19001_003E", "B19001_004E", "B19001_014E", "B19001_015E", "B19001_016E", "B19001_017E", 
       "B17017_002E", "B17017_001E", 
       "C17002_005E", "C17002_004E", "C17002_003E", "C17002_002E", "C17002_001E", 
       "B15003_001E", "B15003_002E", "B15003_003E", "B15003_004E", "B15003_005E", "B15003_006E", "B15003_007E", "B15003_008E", "B15003_009E", "B15003_010E", "B15003_011E", "B15003_012E", "B15003_013E", "B15003_014E", "B15003_015E", "B15003_016E", "B15003_022E", "B15003_023E", "B15003_024E", "B15003_025E",
       "C24010_003E", "C24010_027E", "C24010_039E", "C24010_063E", "C24010_001E",
       "B23025_005E", "B23025_003E",
       "B25077_001E", 
       "B25064_001E", 
       "B25088_002E", 
       "B25003_001E", "B25003_002E",
       "B11005_001E", "B11005_006E", "B11005_007E", 
       "B25044_003E", "B25044_010E", "B25044_001E",
       "B25043_007E", "B25043_016E", "B25043_001E",
       "B25049_007E", "B25049_004E", "B25049_001E", 
       "B25014_005E", "B25014_006E", "B25014_007E", "B25014_011E", "B25014_012E", "B25014_013E", "B25014_001E")
  
# CBG level
ACS_CBG <- data.frame(do.call(rbind, lapply(1:length(states), function(x){
   ACS_CBG <- get_acs(
    year = 2015,
    geography = "block group",
    state = states[x],
    key = Sys.getenv("CENSUS_API_KEY"),
    variables = ADI_Vars)
})))
ACS_CBG <- dcast(ACS_CBG, GEOID + NAME ~ variable, value.var=c("estimate"))
ReADI_CBG_2015_CBG_Raw <- ACS_CBG %>% separate(NAME, c("Block_Group", "Tract", "County", "State"), sep = ",") # 242336
save(ReADI_CBG_2015_CBG_Raw, file = "ReADI_CBG_2015_CBG_Raw.RData")

# census tract level
ACS_CT <- data.frame(do.call(rbind, lapply(1:length(states), function(x){
   ACS_CT <- get_acs(
    year = 2015,
    geography = "tract",
    state = states[x],
    key = Sys.getenv("CENSUS_API_KEY"),
    variables = ADI_Vars)
})))

ACS_CT <- dcast(ACS_CT, GEOID + NAME ~ variable, value.var=c("estimate"))
ReADI_CBG_2015_CT_Raw <- ACS_CT %>% separate(NAME, c("Tract", "County", "State"), sep = ",") # 85396
save(ReADI_CBG_2015_CT_Raw, file = "ReADI_CBG_2015_CT_Raw.RData")

# county level
ACS_C <- data.frame(do.call(rbind, lapply(1:length(states), function(x){
   ACS_C <- get_acs(
    year = 2015,
    geography = "county",
    state = states[x],
    key = Sys.getenv("CENSUS_API_KEY"),
    variables = ADI_Vars)
})))

ACS_C <- dcast(ACS_C, GEOID + NAME ~ variable, value.var=c("estimate"))
ReADI_CBG_2015_C_Raw <- ACS_C %>% separate(NAME, c("County", "State"), sep = ",") # 3222
save(ReADI_CBG_2015_C_Raw, file = "ReADI_CBG_2015_C_Raw.RData")
```
  </div>
</div>
  
<br><br>

### Pulling Geography Measures
***

These values are required for both imputation and for visualizations involving mapping.

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code434">Show Code</button>
  <div id="code434" class="collapse">
```{r, eval = F}
CBG_Geo <- do.call(rbind, lapply(1:length(states), function(x){
   ADI_2015_CBG_Raw_NG <- get_acs(
    year = 2015,
    geography = "block group",
    geometry = TRUE,
    state = states[x],
    key = Sys.getenv("CENSUS_API_KEY"),
    variables = "B01003_001")
}))
CBG_Geo <- CBG_Geo[!CBG_Geo$estimate == 0,] # from 242336 to 240172
ACS_CBG_2015_Geometry <- CBG_Geo[, c("GEOID", "geometry")]
save(ACS_CBG_2015_Geometry, file = "ACS_CBG_2015_Geometry.RData")

CT_Geo <- do.call(rbind, lapply(1:length(states), function(x){
   ADI_2015_CT_Raw_NG <- get_acs(
    year = 2015,
    geography = "tract",
    geometry = TRUE,
    state = states[x],
    key = Sys.getenv("CENSUS_API_KEY"),
    variables = "B01003_001")
}))
CT_Geo <- CT_Geo[!CT_Geo$estimate == 0,] # 85396 to 84539
ACS_CT_2015_Geometry <- CT_Geo[, c("GEOID", "geometry")]
save(ACS_CT_2015_Geometry, file = "ACS_CT_2015_Geometry.RData")

C_Geo <- do.call(rbind, lapply(1:length(states), function(x){
   ADI_2015_C_Raw_NG <- get_acs(
    year = 2015,
    geography = "county",
    geometry = TRUE,
    state = states[x],
    key = Sys.getenv("CENSUS_API_KEY"),
    variables = "B01003_001")
}))
C_Geo <- C_Geo[!C_Geo$estimate == 0,] #3222 as none had 0
ACS_C_2015_Geometry <- C_Geo[, c("GEOID", "geometry")]
save(ACS_C_2015_Geometry, file = "ACS_C_2015_Geometry.RData")
```
  </div>
</div>
  
<br><br>


## Group Quarters
***

There are areas whose populations constitute individuals who are in group quarters which consist of two main groups:
- Institutionalized
  - Adult correctional facilities
  - Nursing facilities
- Non-Institutionalizes
  - College/University student housing
  
Areas whose populations consist totally of group-quartered individuals cannot have an index properly calculated since the relevancy of much of the measures are not applicable and many of the policies and interventions may not be impactful for these populations. The neighborhood ADI addresses these areas by referring to [Roux et al. (2001)](https://www.sciencedirect.com/science/article/pii/S1047279701002216) where they recommend removing areas that had 33% or more of their population living in group quarters (all types).

Unfortunately, the measures of group quarters, specifically tables B26001, B26103, and B26203 have NAs for all three geographies and so we will have to discuss and source out options accordingly.

## Population Size
***

[Roux et al. (2001)](https://www.sciencedirect.com/science/article/pii/S1047279701002216) also recommended removing areas which had a population size < 100 or had less than 30 housing units. This is something to consider although this may remove areas which could benefit from an ADI score. That being said, those with reported populations of 0 should be removed and we will do so below. as there are census tracts with such areas which have to be removed. The amount of areas removed are detailed in the table below.

<div class = "row">
<div class = "col-md-4">
|Area       |Pop = 0   |Remaining |
|:----------|:---------|:---------|
|CBG        |2,239     |218,940   |
|CT         |873       |72,424    |
|County     |0         |3,142     |
</div>

<div class = "col-md-8">
</div>
</div>

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code494">Show Code</button>
  <div id="code494" class="collapse">
```{r, eval = F}
ReADI_CBG_2015_CBG_Raw <- ReADI_CBG_2015_CBG_Raw[!ReADI_CBG_2015_CBG_Raw$B01003_001 == 0,] 
save(ReADI_CBG_2015_CBG_Raw, file = "ReADI_CBG_2015_CBG_Raw.RData")

ReADI_CBG_2015_CT_Raw <- ReADI_CBG_2015_CT_Raw[!ReADI_CBG_2015_CT_Raw$B01003_001 == 0,] 
save(ReADI_CBG_2015_CT_Raw, file = "ReADI_CBG_2015_CT_Raw.RData")

ReADI_CBG_2015_C_Raw <- ReADI_CBG_2015_C_Raw[!ReADI_CBG_2015_C_Raw$B01003_001 == 0,] 
save(ReADI_CBG_2015_C_Raw, file = "ReADI_CBG_2015_C_Raw.RData")
```
  </div>
</div>
  
<br><br>

## Missing Values
***

Below is a summary of the amount of missing values among variables with missing information:

<div class = "row">
<div class = "col-md-6">
|Variable       |CBG N (/%)   |CT N (/%)   |C N (/%)    |
|:--------------|:------------|:-----------|:-----------|
|B19013_001     |5,461 (3)    |367 (1)     |1 (<1)      |
|B25043_001     |380 (<1)     |140 (<1)    |7 (<1)      |
|B25043_007     |380 (<1)     |140 (<1)    |7 (<1)      |
|B25043_016     |380 (<1)     |140 (<1)    |7 (<1)      |
|B25064_001     |38,460 (18)  |1,387 (2)   |1 (<1)      |
|B25077_001     |12,041 (6)   |1,332 (2)   |2 (<1)      |
|B25088_002     |26,916 (12)  |2,241 (3)   |4 (<1)      |
</div>

<div class = "col-md-6">
</div>
</div>

<br>

Among the missing data there were 12 counties which had NAs for one of these variables:

<div style="margin:16px 0">
 <center><button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code569"><h4 class="text-white">County NAs</h4></button></center>
  <div id="code569" class="collapse">
|County, ST          |Population|B19013_001|B25043_001|B25043_007|B25043_016|B25064_001|B25077_001|B25088_002|
|:-------------------|:---------|:---------|:---------|:---------|:---------|:---------|:---------|:---------|
|Lonoke County, AR   |70,691    |\$53,631  |NA        |NA        |NA        |\$714     |\$126,900 |\$1,097   |
|St. Johns County, FL|210,495   |\$66,194  |NA        |NA        |NA        |\$1,119   |\$244,400 |\$1,756   |
|Bartow County, GA   |101,336   |\$48,893  |NA        |NA        |NA        |\$822     |\$123,800 |\$1,173   |
|Catoosa County, GA  |65,375    |\$50,876  |NA        |NA        |NA        |\$730     |\$129,800 |\$1,141   |
|Floyd County, GA    |96,169    |\$41,549  |NA        |NA        |NA        |\$674     |\$117,100 |\$1,133   |
|Kalawao County, HI  |85        |\$66,250  |54        |0         |3         |\$856     |NA        |NA        |
|Campbell County, KY |91,475    |\$54,621  |NA        |NA        |NA        |\$748     |\$150,400 |\$1,241   |
|Storey Parish, NV   |3,929     |\$64,832  |1,767     |28        |0         |NA        |\$177,000 |\$1,348   |
|Kenedy County, TX   |565       |\$36,438  |145       |0         |0         |\$558     |NA        |NA        |
|King County, TX     |267       |\$63,250  |113       |0         |0         |\$743     |\$73,000  |NA        |
|Loving County, TX   |117       |NA        |45        |0         |0         |\$871     |\$92,500  |NA        |
|Portage County, WI  |70,432    |\$51,613  |NA        |NA        |NA        |\$675     |\$151,700 |\$1,216   |
  </div>
</div>

<br>

Of note, the above counties all consist of low population areas which seems to be the source for the missing information. Kalawao County in Hawaii is the most unique in that it is a very isolated and protected community which originated as a quarantine village for those with leprosy and the remaining inhabitants are all that are remaining and will be able to reside there. 

Of important note is of all of the measures which have missing information for all geographical regions, it consists of the four currency variables which are those at the center of the standardization issues. This means that imputation is required only for the very measures which are driving the neighborhood ADI compounding the ADI issues even further. Additionally, there are plumbing variables missing as well but less than 1% would need to be imputed.

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code498">Show Code</button>
  <div id="code498" class="collapse">
```{r, eval = F}
# By Variables
NA_Vars <- ReADI_CBG_2015_CBG_Raw[, colSums(is.na(ReADI_CBG_2015_CBG_Raw)) > 0]
NA_Vars <- as.data.frame(colSums(is.na(NA_Vars))) 
colnames(NA_Vars) <- "CBG_N"

NA_Vars$Perc_NA_CBG <- sapply(1:nrow(NA_Vars), function (x){val <- round((NA_Vars[x,]/nrow(ReADI_CBG_2015_CBG_Raw))*100,2)})
NA_Vars_CT <- ReADI_CBG_2015_CT_Raw[, colSums(is.na(ReADI_CBG_2015_CT_Raw)) > 0]
identical(colnames(NA_Vars_CT), rownames(NA_Vars))
NA_Vars$CT_N <- colSums(is.na(NA_Vars_CT))

NA_Vars$Perc_NA_CT <- sapply(1:nrow(NA_Vars), function (x){val <- round((NA_Vars[x,"CT_N"]/nrow(ReADI_CBG_2015_CT_Raw))*100,2)})
NA_Vars_C <- ReADI_CBG_2015_C_Raw[, colSums(is.na(ReADI_CBG_2015_C_Raw)) > 0]
identical(colnames(NA_Vars_C), rownames(NA_Vars))
NA_Vars$C_N <- colSums(is.na(NA_Vars_C))
NA_Vars$Perc_NA_C <- sapply(1:nrow(NA_Vars), function (x){val <- round((NA_Vars[x,"C_N"]/nrow(ReADI_CBG_2015_C_Raw))*100,2)})

CountiesMissing <- ReADI_CBG_2015_C_Raw[rowSums(is.na(ReADI_CBG_2015_C_Raw)) > 0, c("GEOID", "County", "State", "B01003_001", rownames(NA_Vars))]

# By Geo Area
NA_CBG_GEOs <- ReADI_CBG_2015_CBG_Raw[rowSums(is.na(ReADI_CBG_2015_CBG_Raw)) > 0,]
NA_CBG_GEOs <- as.data.frame(rowSums(is.na(NA_CBG_GEOs))) 
table(NA_CBG_GEOs[1]) # 5 max

NA_CT_GEOs <- ReADI_CBG_2015_CT_Raw[rowSums(is.na(ReADI_CBG_2015_CT_Raw)) > 0,]
NA_CT_GEOs <- as.data.frame(rowSums(is.na(NA_CT_GEOs))) 
table(NA_CT_GEOs[1]) # 4 max

NA_C_GEOs <- ReADI_CBG_2015_C_Raw[rowSums(is.na(ReADI_CBG_2015_C_Raw)) > 0,]
NA_C_GEOs <- as.data.frame(rowSums(is.na(NA_C_GEOs))) 
table(NA_C_GEOs[1]) # 3 max
```
  </div>
</div>
  
<br><br>

### Imputation
***

We will perform a nearest neighbor imputation using type Queen contiguity. The median values of all border touching geographies will be used to replace NAs for the respective geographical levels within state. Imputation will happen iteratively until no further CBGs can be imputed. Among those which can't be imputed in this way we will then impute with the average of the upper level geography if appropriate.

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code531">Imputation Function</button>
  <div id="code531" class="collapse">
```{r, eval = F}
impute.fun <- function(data, var, shape) {
  variable <- data[, c("GEOID", var)]
  colnames(variable) <- c("GEOID", "estimate")
  message(paste0(sum(is.na(variable$estimate)), " missing before imputation."))
  df <- shape %>% left_join(variable, by = "GEOID")
  df <- df[!is.na(st_dimension(df)),]
  df$State <- substr(df$GEOID, 1, 2)
  df$raw.estimate <- df$estimate
  
  # set check values for iterative while loop:
  missing.pre <- 1
  missing.post <- 0

  # while loop for iterations!
  while(missing.pre > missing.post) { 
    missing.pre <- sum(is.na(df$estimate))
    imputed <- bind_rows(lapply(unique(df$State), function(x){
      df2 <- df[df$State == x,]
      df2$index <- 1:nrow(df2)
      df2$nb <- spdep::poly2nb(df2, queen = TRUE)
      df2$geometry <- NULL
      df2$imputed <- unlist(lapply(df2$nb, function(z){
        median(df2$estimate[z], na.rm = T)}))
      df2 <- df2[, c("GEOID", "estimate", "State", "imputed")]
      return(df2)
      }))
    df$imputed <- NULL
    df <- left_join(df, imputed, by = c("GEOID", "estimate", "State"))
    df$estimate <- ifelse(is.na(df$estimate), df$imputed, df$estimate)
    missing.post <- sum(is.na(df$estimate))
  }
  message(paste0(sum(is.na(df$estimate)), " missing after imputing."))
  return(df)
}
```
  </div>
</div>
  
<br><br>

> CBG

Imputation was successful in that only the NAs were imputed and replaced in the overall dataset though there are some NAs remaining which will be investigated below. Out of interest the following table will include how correlated the actual measure is to a fully imputed measure.

<div class = "row">
<div class = "col-md-6">
|Variable       |$r$         |*p*value        |NAs Remaining   |
|:--------------|:-----------|:---------------|:---------------|
|B19013_001     |0.78        |$<2.2x10^{-16}$ |10              |
|B25043_001     |0.42        |$<2.2x10^{-16}$ |0               |
|B25043_007     |0.24        |$<2.2x10^{-16}$ |0               |
|B25043_016     |0.29        |$<2.2x10^{-16}$ |0               |
|B25064_001     |0.80        |$<2.2x10^{-16}$ |16              |
|B25077_001     |0.91        |$<2.2x10^{-16}$ |8               |
|B25088_002     |0.89        |$<2.2x10^{-16}$ |10              |
</div>

<div class = "col-md-6">
</div>
</div>

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code581">Show Code</button>
  <div id="code581" class="collapse">
```{r, eval = F}
load("ACS_CBG_2015_Geometry.RData")
load("ReADI_CBG_2015_CBG_Raw.RData")

B19013_Impute_CBG <- impute.fun(data = ReADI_CBG_2015_CBG_Raw, var = "B19013_001", shape = ACS_CBG_2015_Geometry) 
head(B19013_Impute_CBG)
cor.test(B19013_Impute_CBG$estimate, B19013_Impute_CBG$raw.estimate) # should always be 1 as NAs are removed to ensure nothing changed to the non-missing data.
cor.test(B19013_Impute_CBG$imputed, B19013_Impute_CBG$raw.estimate) 
B19013_Impute_CBG$geometry <- NULL
B19013_Impute_CBG <- B19013_Impute_CBG[, c("GEOID", "estimate")]
colnames(B19013_Impute_CBG) <- c("GEOID", "B19013_001")

B25043_001_Impute_CBG <- impute.fun(data = ReADI_CBG_2015_CBG_Raw, var = "B25043_001", shape = ACS_CBG_2015_Geometry)
cor.test(B25043_001_Impute_CBG$estimate, B25043_001_Impute_CBG$raw.estimate)
cor.test(B25043_001_Impute_CBG$imputed, B25043_001_Impute_CBG$raw.estimate) 
B25043_001_Impute_CBG$geometry <- NULL
B25043_001_Impute_CBG <- B25043_001_Impute_CBG[, c("GEOID", "estimate")]
colnames(B25043_001_Impute_CBG) <- c("GEOID", "B25043_001")

B25043_007_Impute_CBG <- impute.fun(data = ReADI_CBG_2015_CBG_Raw, var = "B25043_007", shape = ACS_CBG_2015_Geometry)
cor.test(B25043_007_Impute_CBG$estimate, B25043_007_Impute_CBG$raw.estimate)
cor.test(B25043_007_Impute_CBG$imputed, B25043_007_Impute_CBG$raw.estimate) 
B25043_007_Impute_CBG$geometry <- NULL
B25043_007_Impute_CBG <- B25043_007_Impute_CBG[, c("GEOID", "estimate")]
colnames(B25043_007_Impute_CBG) <- c("GEOID", "B25043_007")

B25043_016_Impute_CBG <- impute.fun(data = ReADI_CBG_2015_CBG_Raw, var = "B25043_016", shape = ACS_CBG_2015_Geometry)
cor.test(B25043_016_Impute_CBG$estimate, B25043_016_Impute_CBG$raw.estimate)
cor.test(B25043_016_Impute_CBG$imputed, B25043_016_Impute_CBG$raw.estimate) 
B25043_016_Impute_CBG$geometry <- NULL
B25043_016_Impute_CBG <- B25043_016_Impute_CBG[, c("GEOID", "estimate")]
colnames(B25043_016_Impute_CBG) <- c("GEOID", "B25043_016")

B25064_Impute_CBG <- impute.fun(data = ReADI_CBG_2015_CBG_Raw, var = "B25064_001", shape = ACS_CBG_2015_Geometry)
cor.test(B25064_Impute_CBG$estimate, B25064_Impute_CBG$raw.estimate)
cor.test(B25064_Impute_CBG$imputed, B25064_Impute_CBG$raw.estimate) 
B25064_Impute_CBG$geometry <- NULL
B25064_Impute_CBG <- B25064_Impute_CBG[, c("GEOID", "estimate")]
colnames(B25064_Impute_CBG) <- c("GEOID", "B25064_001")

B25077_Impute_CBG <- impute.fun(data = ReADI_CBG_2015_CBG_Raw, var = "B25077_001", shape = ACS_CBG_2015_Geometry)
cor.test(B25077_Impute_CBG$estimate, B25077_Impute_CBG$raw.estimate)
cor.test(B25077_Impute_CBG$imputed, B25077_Impute_CBG$raw.estimate) 
B25077_Impute_CBG$geometry <- NULL
B25077_Impute_CBG <- B25077_Impute_CBG[, c("GEOID", "estimate")]
colnames(B25077_Impute_CBG) <- c("GEOID", "B25077_001")

B25088_Impute_CBG <- impute.fun(data = ReADI_CBG_2015_CBG_Raw, var = "B25088_002", shape = ACS_CBG_2015_Geometry)
cor.test(B25088_Impute_CBG$estimate, B25088_Impute_CBG$raw.estimate)
cor.test(B25088_Impute_CBG$imputed, B25088_Impute_CBG$raw.estimate) 
B25088_Impute_CBG$geometry <- NULL
B25088_Impute_CBG <- B25088_Impute_CBG[, c("GEOID", "estimate")]
colnames(B25088_Impute_CBG) <- c("GEOID", "B25088_002")

CBG_2015_Imputed <- plyr::join(B19013_Impute_CBG, B25043_001_Impute_CBG, by = "GEOID")
CBG_2015_Imputed <- plyr::join(CBG_2015_Imputed, B25043_007_Impute_CBG, by = "GEOID")
CBG_2015_Imputed <- plyr::join(CBG_2015_Imputed, B25043_016_Impute_CBG, by = "GEOID")
CBG_2015_Imputed <- plyr::join(CBG_2015_Imputed, B25064_Impute_CBG, by = "GEOID")
CBG_2015_Imputed <- plyr::join(CBG_2015_Imputed, B25077_Impute_CBG, by = "GEOID")
CBG_2015_Imputed <- plyr::join(CBG_2015_Imputed, B25088_Impute_CBG, by = "GEOID")
save(CBG_2015_Imputed, file = "CBG_2015_Imputed.RData")

sum(is.na(CBG_2015_Imputed$B19013_001)) #10
sum(is.na(CBG_2015_Imputed$B25043_001)) #0
sum(is.na(CBG_2015_Imputed$B25043_007)) #0
sum(is.na(CBG_2015_Imputed$B25043_016)) #0
sum(is.na(CBG_2015_Imputed$B25064_001)) #16
sum(is.na(CBG_2015_Imputed$B25077_001)) #8
sum(is.na(CBG_2015_Imputed$B25088_002)) #10

ReADI_CBG_2015_Imp <- ReADI_CBG_2015_CBG_Raw[, !colnames(ReADI_CBG_2015_CBG_Raw) %in% c("B19013_001", "B25043_001", "B25043_007", "B25043_016", "B25064_001", "B25077_001", "B25088_002")]
ReADI_CBG_2015_Imp <- plyr::join(ReADI_CBG_2015_Imp, CBG_2015_Imputed, by = "GEOID")
save(ReADI_CBG_2015_Imp, file = "ReADI_CBG_2015_Imp.RData")
```
  </div>
</div>
  
<br><br>

> CT

<div class = "row">
<div class = "col-md-6">
|Variable       |$r$         |*p*value        |NAs Remaining   |
|:--------------|:-----------|:---------------|:---------------|
|B19013_001     |0.81        |$<2.2x10^{-16}$ |3               |
|B25043_001     |0.41        |$<2.2x10^{-16}$ |0               |
|B25043_007     |0.38        |$<2.2x10^{-16}$ |0               |
|B25043_016     |0.44        |$<2.2x10^{-16}$ |0               |
|B25064_001     |0.81        |$<2.2x10^{-16}$ |3               |
|B25077_001     |0.91        |$<2.2x10^{-16}$ |3               |
|B25088_002     |0.90        |$<2.2x10^{-16}$ |3               |
</div>

<div class = "col-md-6">
</div>
</div>

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code639">Show Code</button>
  <div id="code639" class="collapse">
```{r, eval = F}
load("ACS_CT_2015_Geometry.RData")
load("ReADI_CBG_2015_CT_Raw.RData")

B19013_Impute_CT <- impute.fun(data = ReADI_CBG_2015_CT_Raw, var = "B19013_001", shape = ACS_CT_2015_Geometry)
cor.test(B19013_Impute_CT$estimate, B19013_Impute_CT$raw.estimate) # should always be 1, as nas are removed, to ensure nothing changed to the non-missing data.
cor.test(B19013_Impute_CT$imputed, B19013_Impute_CT$raw.estimate) 
B19013_Impute_CT$geometry <- NULL
B19013_Impute_CT <- B19013_Impute_CT[, c("GEOID", "estimate")]
colnames(B19013_Impute_CT) <- c("GEOID", "B19013_001")

B25043_1_Impute_CT <- impute.fun(data = ReADI_CBG_2015_CT_Raw, var = "B25043_001", shape = ACS_CT_2015_Geometry)
cor.test(B25043_1_Impute_CT$estimate, B25043_1_Impute_CT$raw.estimate)
cor.test(B25043_1_Impute_CT$imputed, B25043_1_Impute_CT$raw.estimate) # 0.41 p < 2.2e-16 - this is low BUT only 140 are replaced.
B25043_1_Impute_CT$geometry <- NULL
B25043_1_Impute_CT <- B25043_1_Impute_CT[, c("GEOID", "estimate")]
colnames(B25043_1_Impute_CT) <- c("GEOID", "B25043_001")

B25043_7_Impute_CT <- impute.fun(data = ReADI_CBG_2015_CT_Raw, var = "B25043_007", shape = ACS_CT_2015_Geometry)
cor.test(B25043_7_Impute_CT$estimate, B25043_7_Impute_CT$raw.estimate)
cor.test(B25043_7_Impute_CT$imputed, B25043_7_Impute_CT$raw.estimate) # 0.38 p < 2.2e-16 - this is low BUT only 140 are replaced.
B25043_7_Impute_CT$geometry <- NULL
B25043_7_Impute_CT <- B25043_7_Impute_CT[, c("GEOID", "estimate")]
colnames(B25043_7_Impute_CT) <- c("GEOID", "B25043_007")

B25043_16_Impute_CT <- impute.fun(data = ReADI_CBG_2015_CT_Raw, var = "B25043_016", shape = ACS_CT_2015_Geometry)
cor.test(B25043_16_Impute_CT$estimate, B25043_16_Impute_CT$raw.estimate)
cor.test(B25043_16_Impute_CT$imputed, B25043_16_Impute_CT$raw.estimate) # 0.44 p < 2.2e-16 - this is low BUT only 140 are replaced.
B25043_16_Impute_CT$geometry <- NULL
B25043_16_Impute_CT <- B25043_16_Impute_CT[, c("GEOID", "estimate")]
colnames(B25043_16_Impute_CT) <- c("GEOID", "B25043_016")

B25064_Impute_CT <- impute.fun(data = ReADI_CBG_2015_CT_Raw, var = "B25064_001", shape = ACS_CT_2015_Geometry)
cor.test(B25064_Impute_CT$estimate, B25064_Impute_CT$raw.estimate)
cor.test(B25064_Impute_CT$imputed, B25064_Impute_CT$raw.estimate) 
B25064_Impute_CT$geometry <- NULL
B25064_Impute_CT <- B25064_Impute_CT[, c("GEOID", "estimate")]
colnames(B25064_Impute_CT) <- c("GEOID", "B25064_001")

B25077_Impute_CT <- impute.fun(data = ReADI_CBG_2015_CT_Raw, var = "B25077_001", shape = ACS_CT_2015_Geometry)
cor.test(B25077_Impute_CT$estimate, B25077_Impute_CT$raw.estimate)
cor.test(B25077_Impute_CT$imputed, B25077_Impute_CT$raw.estimate) 
B25077_Impute_CT$geometry <- NULL
B25077_Impute_CT <- B25077_Impute_CT[, c("GEOID", "estimate")]
colnames(B25077_Impute_CT) <- c("GEOID", "B25077_001")

B25088_Impute_CT <- impute.fun(data = ReADI_CBG_2015_CT_Raw, var = "B25088_002", shape = ACS_CT_2015_Geometry)
cor.test(B25088_Impute_CT$estimate, B25088_Impute_CT$raw.estimate)
cor.test(B25088_Impute_CT$imputed, B25088_Impute_CT$raw.estimate) 
B25088_Impute_CT$geometry <- NULL
B25088_Impute_CT <- B25088_Impute_CT[, c("GEOID", "estimate")]
colnames(B25088_Impute_CT) <- c("GEOID", "B25088_002")

CT_2015_Imputed <- plyr::join(B19013_Impute_CT, B25043_1_Impute_CT, by = "GEOID")
CT_2015_Imputed <- plyr::join(CT_2015_Imputed, B25043_7_Impute_CT, by = "GEOID")
CT_2015_Imputed <- plyr::join(CT_2015_Imputed, B25043_16_Impute_CT, by = "GEOID")
CT_2015_Imputed <- plyr::join(CT_2015_Imputed, B25064_Impute_CT, by = "GEOID")
CT_2015_Imputed <- plyr::join(CT_2015_Imputed, B25077_Impute_CT, by = "GEOID")
CT_2015_Imputed <- plyr::join(CT_2015_Imputed, B25088_Impute_CT, by = "GEOID")
save(CT_2015_Imputed, file = "CT_2015_Imputed.RData")

sum(is.na(CT_2015_Imputed$B19013_001)) #3
sum(is.na(CT_2015_Imputed$B25043_001)) #0
sum(is.na(CT_2015_Imputed$B25043_007)) #0
sum(is.na(CT_2015_Imputed$B25043_016)) #0
sum(is.na(CT_2015_Imputed$B25064_001)) #3
sum(is.na(CT_2015_Imputed$B25077_001)) #3
sum(is.na(CT_2015_Imputed$B25088_002)) #3 

ReADI_CT_2015_Imp <- ReADI_CBG_2015_CT_Raw[, !colnames(ReADI_CBG_2015_CT_Raw) %in% c("B19013_001", "B25043_001", "B25043_007", "B25043_016", "B25064_001", "B25077_001", "B25088_002")]
ReADI_CT_2015_Imp <- plyr::join(ReADI_CT_2015_Imp, CT_2015_Imputed, by = "GEOID")
save(ReADI_CT_2015_Imp, file = "ReADI_CT_2015_Imp.RData")
```
  </div>
</div>
  
<br><br>

> C

<div class = "row">
<div class = "col-md-6">
|Variable       |$r$          |*p*value       |NAs Remaining   |
|:--------------|:------------|:--------------|:---------------|
|B19013_001     |0.78        |$<2.2x10^{-16}$ |0               |
|B25043_001     |0.54        |$<2.2x10^{-16}$ |0               |
|B25043_007     |0.49        |$<2.2x10^{-16}$ |0               |
|B25043_016     |0.48        |$<2.2x10^{-16}$ |0               |
|B25064_001     |0.84        |$<2.2x10^{-16}$ |0               |
|B25077_001     |0.85        |$<2.2x10^{-16}$ |0               |
|B25088_002     |0.87        |$<2.2x10^{-16}$ |0               |
</div>

<div class = "col-md-6">
</div>
</div>

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code701">Show Code</button>
  <div id="code701" class="collapse">
```{r, eval = F}
load("ACS_C_2015_Geometry.RData")
load("ReADI_CBG_2015_C_Raw.RData")

## For the code to work you need to remove state 11 (DC) as it has only one county and will trigger an error.
table(ReADI_CBG_2015_C_Raw$State)
data = ReADI_CBG_2015_C_Raw[!ReADI_CBG_2015_C_Raw$State == " District of Columbia",]
geoid_rm <- ReADI_CBG_2015_C_Raw[ReADI_CBG_2015_C_Raw$State == " District of Columbia", "GEOID"]
shape = ACS_C_2015_Geometry[!ACS_C_2015_Geometry$GEOID %in% geoid_rm,]

B19013_Impute_C <- impute.fun(data = data, var = "B19013_001", shape = shape)
cor.test(B19013_Impute_C$estimate, B19013_Impute_C$raw.estimate)
cor.test(B19013_Impute_C$imputed, B19013_Impute_C$raw.estimate) 
B19013_Impute_C$geometry <- NULL
B19013_Impute_C <- B19013_Impute_C[, c("GEOID", "estimate")]
colnames(B19013_Impute_C) <- c("GEOID", "B19013_001")

B25043_1_Impute_C <- impute.fun(data = data, var = "B25043_001", shape = shape)
cor.test(B25043_1_Impute_C$estimate, B25043_1_Impute_C$raw.estimate)
cor.test(B25043_1_Impute_C$imputed, B25043_1_Impute_C$raw.estimate) # 0.54 p < 2.2e-16 - this is low BUT only 7 are replaced.
B25043_1_Impute_C$geometry <- NULL
B25043_1_Impute_C <- B25043_1_Impute_C[, c("GEOID", "estimate")]
colnames(B25043_1_Impute_C) <- c("GEOID", "B25043_001")

B25043_7_Impute_C <- impute.fun(data = data, var = "B25043_007", shape = shape)
cor.test(B25043_7_Impute_C$estimate, B25043_7_Impute_C$raw.estimate)
cor.test(B25043_7_Impute_C$imputed, B25043_7_Impute_C$raw.estimate) # 0.49 p < 2.2e-16 - this is low BUT only 7 are replaced.
B25043_7_Impute_C$geometry <- NULL
B25043_7_Impute_C <- B25043_7_Impute_C[, c("GEOID", "estimate")]
colnames(B25043_7_Impute_C) <- c("GEOID", "B25043_007")

B25043_16_Impute_C <- impute.fun(data = data, var = "B25043_016", shape = shape)
cor.test(B25043_16_Impute_C$estimate, B25043_16_Impute_C$raw.estimate)
cor.test(B25043_16_Impute_C$imputed, B25043_16_Impute_C$raw.estimate) # 0.48 p < 2.2e-16 - this is low BUT only 7 are replaced.
B25043_16_Impute_C$geometry <- NULL
B25043_16_Impute_C <- B25043_16_Impute_C[, c("GEOID", "estimate")]
colnames(B25043_16_Impute_C) <- c("GEOID", "B25043_016")

B25064_Impute_C <- impute.fun(data = data, var = "B25064_001", shape = shape)
cor.test(B25064_Impute_C$estimate, B25064_Impute_C$raw.estimate)
cor.test(B25064_Impute_C$imputed, B25064_Impute_C$raw.estimate) 
B25064_Impute_C$geometry <- NULL
B25064_Impute_C <- B25064_Impute_C[, c("GEOID", "estimate")]
colnames(B25064_Impute_C) <- c("GEOID", "B25064_001")

B25077_Impute_C <- impute.fun(data = data, var = "B25077_001", shape = shape)
cor.test(B25077_Impute_C$estimate, B25077_Impute_C$raw.estimate)
cor.test(B25077_Impute_C$imputed, B25077_Impute_C$raw.estimate) 
B25077_Impute_C$geometry <- NULL
B25077_Impute_C <- B25077_Impute_C[, c("GEOID", "estimate")]
colnames(B25077_Impute_C) <- c("GEOID", "B25077_001")

B25088_Impute_C <- impute.fun(data = data, var = "B25088_002", shape = shape)
cor.test(B25088_Impute_C$estimate, B25088_Impute_C$raw.estimate)
cor.test(B25088_Impute_C$imputed, B25088_Impute_C$raw.estimate) 
B25088_Impute_C$geometry <- NULL
B25088_Impute_C <- B25088_Impute_C[, c("GEOID", "estimate")]
colnames(B25088_Impute_C) <- c("GEOID", "B25088_002")

C_2015_Imputed <- plyr::join(B19013_Impute_C, B25043_1_Impute_C, by = "GEOID")
C_2015_Imputed <- plyr::join(C_2015_Imputed, B25043_7_Impute_C, by = "GEOID")
C_2015_Imputed <- plyr::join(C_2015_Imputed, B25043_16_Impute_C, by = "GEOID")
C_2015_Imputed <- plyr::join(C_2015_Imputed, B25064_Impute_C, by = "GEOID")
C_2015_Imputed <- plyr::join(C_2015_Imputed, B25077_Impute_C, by = "GEOID")
C_2015_Imputed <- plyr::join(C_2015_Imputed, B25088_Impute_C, by = "GEOID")
save(C_2015_Imputed, file = "C_2015_Imputed.RData")

sum(is.na(C_2015_Imputed$B19013_001)) #0
sum(is.na(C_2015_Imputed$B25043_001)) #0
sum(is.na(C_2015_Imputed$B25043_007)) #0
sum(is.na(C_2015_Imputed$B25043_016)) #0
sum(is.na(C_2015_Imputed$B25064_001)) #0
sum(is.na(C_2015_Imputed$B25077_001)) #0
sum(is.na(C_2015_Imputed$B25088_002)) #0

ReADI_C_2015_Imp <- ReADI_CBG_2015_C_Raw[, !colnames(ReADI_CBG_2015_C_Raw) %in% c("B19013_001", "B25043_001", "B25043_007", "B25043_016", "B25064_001", "B25077_001", "B25088_002")]
ReADI_C_2015_Imp <- plyr::join(ReADI_C_2015_Imp, C_2015_Imputed, by = "GEOID")

# replacing District of Columbia data removed in previous line.
identical(ReADI_CBG_2015_C_Raw$GEOID, ReADI_C_2015_Imp$GEOID)
ReADI_C_2015_Imp$B19013_001 <- ifelse(is.na(ReADI_C_2015_Imp$B19013_001), ReADI_CBG_2015_C_Raw$B19013_001, ReADI_C_2015_Imp$B19013_001)
ReADI_C_2015_Imp$B25043_001 <- ifelse(is.na(ReADI_C_2015_Imp$B25043_001), ReADI_CBG_2015_C_Raw$B25043_001, ReADI_C_2015_Imp$B25043_001)
ReADI_C_2015_Imp$B25043_007 <- ifelse(is.na(ReADI_C_2015_Imp$B25043_007), ReADI_CBG_2015_C_Raw$B25043_007, ReADI_C_2015_Imp$B25043_007)
ReADI_C_2015_Imp$B25043_016 <- ifelse(is.na(ReADI_C_2015_Imp$B25043_016), ReADI_CBG_2015_C_Raw$B25043_016, ReADI_C_2015_Imp$B25043_016)
ReADI_C_2015_Imp$B25064_001 <- ifelse(is.na(ReADI_C_2015_Imp$B25064_001), ReADI_CBG_2015_C_Raw$B25064_001, ReADI_C_2015_Imp$B25064_001)
ReADI_C_2015_Imp$B25077_001 <- ifelse(is.na(ReADI_C_2015_Imp$B25077_001), ReADI_CBG_2015_C_Raw$B25077_001, ReADI_C_2015_Imp$B25077_001)
ReADI_C_2015_Imp$B25088_002 <- ifelse(is.na(ReADI_C_2015_Imp$B25088_002), ReADI_CBG_2015_C_Raw$B25088_002, ReADI_C_2015_Imp$B25088_002)
save(ReADI_C_2015_Imp, file = "ReADI_C_2015_Imp.RData")
```
  </div>
</div>
  
<br><br>

### Missing After Imputation
***

<br>

The plumbing variables were all successfully imputed. The currency variables have some missing values remaining. The main reason why there would still be missing information after imputation is due to the imputation method used which required measures due to bordering geographies which would not impute areas which have no bordering areas such as islands. We will investigate what areas are remaining and if it makes sense to impute for these regions.

> Census Tracts

Of the missing data remaining after imputation there are 9 total census tracts, 8 of which are NAs for all measures. Information based on [Census Reporter](https://censusreporter.org/profiles/):

<div style="margin:16px 0">
 <center><button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code855"><h4 class="text-white">CT NAs</h4></button></center>
  <div id="code855" class="collapse">

|GEOID      |CT   |Location                  |Population|CT Information                         |
|:----------|:----|:-------------------------|:---------|:--------------------------------------|
|06111980000|9800 |Ventura County, CA        |110       |Naval Facility San Nicolas Island.     |
|12087980100|9801 |Monroe County, FL         |8         |Dry Tortugas is a National Park.       |
|26083980100|9801 |Keweenaw County, MI       |15        |Isle Royale Island is a National Park. |
  </div>
</div>

As all of the census tracts which have remaining NAs, they are either national parks whose population consist of rangers or a naval base these CTs will be removed.

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code787">Show Code</button>
  <div id="code787" class="collapse">
```{r, eval = F}
load("ReADI_CT_2015_Imp.RData")
load("ACS_CT_2015_Geometry.RData")
B19013_001_NAs <- ReADI_CT_2015_Imp[is.na(ReADI_CT_2015_Imp$B19013_001), c("GEOID", "Tract", "County", "State", "B01003_001")] 
B25064_001_NAs <- ReADI_CT_2015_Imp[is.na(ReADI_CT_2015_Imp$B25064_001), c("GEOID", "Tract", "County", "State", "B01003_001")] 
B25077_001_NAs <- ReADI_CT_2015_Imp[is.na(ReADI_CT_2015_Imp$B25077_001), c("GEOID", "Tract", "County", "State", "B01003_001")] 
B25088_002_NAs <- ReADI_CT_2015_Imp[is.na(ReADI_CT_2015_Imp$B25088_002), c("GEOID", "Tract", "County", "State", "B01003_001")] 
CT_Missing <- merge(B19013_001_NAs, B25064_001_NAs, all = T)
CT_Missing <- merge(CT_Missing, B25077_001_NAs, all = T)
CT_Missing <- merge(CT_Missing, B25088_002_NAs, all = T)
CT_rm <- c("06111980000", "12087980100", "26083980100")
ReADI_CT_2015_Imp <- ReADI_CT_2015_Imp[!ReADI_CT_2015_Imp$GEOID %in% CT_rm, ] # 72421 remaining - 4 removed
save(ReADI_CT_2015_Imp, file = "ReADI_CT_2015_Imp.RData")
```
  </div>
</div>
  
<br><br>

> Census Block Groups

Observing the specific census block groups which remain NAs after imputation summarized in the table below: 


<div style="margin:16px 0">
 <center><button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code855"><h4 class="text-white">CBG NAs></h4></button></center>
  <div id="code855" class="collapse">
|GEOID       |CBG.CT   |Location                  |Population|CT Information                        |Removed|
|:-----------|:--------|:-------------------------|:---------|:-------------------------------------|:------|
|060375991001|1.5991   |Los Angeles County, CA    |158       |San Clemente Islands                  |       |
|061119800001|1.9800   |Ventura County, CA        |110       |Naval Facility San Nicolas Island     |<strong><strong>X</strong></strong>|
|120879801001|1.9801   |Monroe County, FL         |8         |Dry Tortugas is a National Park       |<strong><strong>X</strong></strong>|
|230139711004|4.9711   |Knox County, ME.          |124       |Isle au Haut and Matinicus Isle       |       |
|230159756005|5.9756   |Lincoln County, ME        |34        |Monhegan Island                       |       |
|250072004006|6.2004   |Dukes County, MA          |75        |Gosnold                               |       |
|260839801001|1.9801   |Keweenaw County, MI       |15        |Isle Royale Island is a National Park |<strong><strong>X</strong></strong>|
|260979505001|1.9505   |Mackinac County, MI       |78        |Bois Blanc Island                     |       |
|360050001001|1.1      |Bronx County, NY          |7,703     |Rikers Island Prison                  |<strong><strong>X</strong></strong>|
|360610240001|1.240    |New York County, NY       |1,688     |                                      |       |
|390430419004|4.419    |Erie County, OH           |138       |Kelleys Island                        |       |
|440050401031|1.401.03 |Newport County, RI        |108       |                                      |       |
|440050401033|3.401.03 |Newport County, RI        |234       |                                      |       |
|440050411002|2.411    |Newport County, RI        |51        |                                      |       |
|510010906001|1.906    |Accomack County, VA       |491       |Tangier Island                        |       |
|530530726035|5.726.03 |Pierce County, WA         |419       |McNeil Island                         |       |
  </div>
</div>

Of the CBGs which remain missing after imputation it is clear that the vast majority of them were not imputed due to simply not directly bordering any other CBGs. These will be given the values attributed to the CT they are a part of for imputation. There are a few which will be removed as indicated in the removed column above based on if they represent national parks containing small population of rangers and areas consisting fully of group quarters.

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code860">Show Code</button>
  <div id="code860" class="collapse">
```{r, eval = F}
load("ReADI_CBG_2015_Imp.RData")
load("ReADI_CT_2015_Imp.RData")
B19013_001_NAs <- ReADI_CBG_2015_Imp[is.na(ReADI_CBG_2015_Imp$B19013_001), c("GEOID", "Tract", "County", "State", "B01003_001")] # 10
B25064_001_NAs <- ReADI_CBG_2015_Imp[is.na(ReADI_CBG_2015_Imp$B25064_001), c("GEOID", "Tract", "County", "State", "B01003_001")] # 16
B25077_001_NAs <- ReADI_CBG_2015_Imp[is.na(ReADI_CBG_2015_Imp$B25077_001), c("GEOID", "Tract", "County", "State", "B01003_001")] # 8
B25088_002_NAs <- ReADI_CBG_2015_Imp[is.na(ReADI_CBG_2015_Imp$B25088_002), c("GEOID", "Tract", "County", "State", "B01003_001")] # 10
CBG_Missing <- merge(B19013_001_NAs, B25064_001_NAs, all = T)
CBG_Missing <- merge(CBG_Missing, B25077_001_NAs, all = T)
CBG_Missing <- merge(CBG_Missing, B25088_002_NAs, all = T)
CBG_Missing <- unique(CBG_Missing)
CBG_rm <- c("061119800001", "120879801001", "260839801001", "360050001001")
ReADI_CBG_2015_Imp <- ReADI_CBG_2015_Imp[!ReADI_CBG_2015_Imp$GEOID %in% CBG_rm, ] # 216701 to 216697 remaining - 4 removed
```
  </div>
</div>
  
<br><br>

### Imputing Remaining CBGs
***

The remaining CBGs will be imputed with the CT measure they are located in resulting in the final results:

<div class = "row">
<div class = "col-md-4">
|Area   |NAs |Amount  |
|:------|:---|:-------|
|CBG    |0   |216,697 |
|CT     |0   |72,421  |
|C      |0   |3,142   |

</div>

<div class = "col-md-8">
</div>
</div>

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code1019">Show Code</button>
  <div id="code1019" class="collapse">
```{r, eval = F}
impute_CBG_geo <- CBG_Missing[!CBG_Missing$GEOID %in% CBG_rm, "GEOID"]
impute_CT_geo <- sapply(length(impute_CBG_geo), function(x){
  substring(impute_CBG_geo, 1, 11)
})
impute_CT_geo <- impute_CT_geo[,1]

impute_data <- ReADI_CT_2015_Imp[ReADI_CT_2015_Imp$GEOID %in% impute_CT_geo, c("GEOID", "B19013_001", "B25064_001", "B25077_001", "B25088_002")]
ReADI_CBG_2015_Imp$GEOID_CT <- substring(ReADI_CBG_2015_Imp$GEOID, 1, 11)
ReADI_CBG_2015_Imp$B19013_001_T <- ReADI_CBG_2015_Imp$B19013_001
ReADI_CBG_2015_Imp$B25064_001_T <- ReADI_CBG_2015_Imp$B25064_001
ReADI_CBG_2015_Imp$B25077_001_T <- ReADI_CBG_2015_Imp$B25077_001
ReADI_CBG_2015_Imp$B25088_002_T <- ReADI_CBG_2015_Imp$B25088_002

# B19013_001
CBG_Imp <- as.data.frame(sapply(1:nrow(impute_data), function(x){
  ifelse(ReADI_CBG_2015_Imp$GEOID_CT == impute_data[x, "GEOID"] & is.na(ReADI_CBG_2015_Imp[,"B19013_001_T"]), impute_data[x,"B19013_001"], ReADI_CBG_2015_Imp$B19013_001_T)
}))
ReADI_CBG_2015_Imp$B19013_001 <- coalesce(CBG_Imp[,1],CBG_Imp[,2],CBG_Imp[,3],CBG_Imp[,4],CBG_Imp[,5],CBG_Imp[,6],CBG_Imp[,7],CBG_Imp[,8],CBG_Imp[,9],CBG_Imp[,10],CBG_Imp[,11])
cor.test(ReADI_CBG_2015_Imp$B19013_001_T, ReADI_CBG_2015_Imp$B19013_001)
sum(is.na(ReADI_CBG_2015_Imp$B19013_001_T))
sum(is.na(ReADI_CBG_2015_Imp$B19013_001))

# B25064_001
CBG_Imp <- as.data.frame(sapply(1:nrow(impute_data), function(x){
  ifelse(ReADI_CBG_2015_Imp$GEOID_CT == impute_data[x, "GEOID"] & is.na(ReADI_CBG_2015_Imp[,"B25064_001_T"]), impute_data[x,"B25064_001"], ReADI_CBG_2015_Imp$B25064_001_T)
}))
ReADI_CBG_2015_Imp$B25064_001 <- coalesce(CBG_Imp[,1],CBG_Imp[,2],CBG_Imp[,3],CBG_Imp[,4],CBG_Imp[,5],CBG_Imp[,6],CBG_Imp[,7],CBG_Imp[,8],CBG_Imp[,9],CBG_Imp[,10],CBG_Imp[,11])
cor.test(ReADI_CBG_2015_Imp$B25064_001_T, ReADI_CBG_2015_Imp$B25064_001)
sum(is.na(ReADI_CBG_2015_Imp$B25064_001_T))
sum(is.na(ReADI_CBG_2015_Imp$B25064_001))

# B25077_001
CBG_Imp <- as.data.frame(sapply(1:nrow(impute_data), function(x){
  ifelse(ReADI_CBG_2015_Imp$GEOID_CT == impute_data[x, "GEOID"] & is.na(ReADI_CBG_2015_Imp[,"B25077_001_T"]), impute_data[x,"B25077_001"], ReADI_CBG_2015_Imp$B25077_001_T)
}))
ReADI_CBG_2015_Imp$B25077_001 <- coalesce(CBG_Imp[,1],CBG_Imp[,2],CBG_Imp[,3],CBG_Imp[,4],CBG_Imp[,5],CBG_Imp[,6],CBG_Imp[,7],CBG_Imp[,8],CBG_Imp[,9],CBG_Imp[,10],CBG_Imp[,11])
cor.test(ReADI_CBG_2015_Imp$B25077_001_T, ReADI_CBG_2015_Imp$B25077_001)
sum(is.na(ReADI_CBG_2015_Imp$B25077_001_T))
sum(is.na(ReADI_CBG_2015_Imp$B25077_001))

# B25088_002
CBG_Imp <- as.data.frame(sapply(1:nrow(impute_data), function(x){
  ifelse(ReADI_CBG_2015_Imp$GEOID_CT == impute_data[x, "GEOID"] & is.na(ReADI_CBG_2015_Imp[,"B25088_002_T"]), impute_data[x,"B25088_002"], ReADI_CBG_2015_Imp$B25088_002_T)
}))
ReADI_CBG_2015_Imp$B25088_002 <- coalesce(CBG_Imp[,1],CBG_Imp[,2],CBG_Imp[,3],CBG_Imp[,4],CBG_Imp[,5],CBG_Imp[,6],CBG_Imp[,7],CBG_Imp[,8],CBG_Imp[,9],CBG_Imp[,10],CBG_Imp[,11])
cor.test(ReADI_CBG_2015_Imp$B25088_002_T, ReADI_CBG_2015_Imp$B25088_002)
sum(is.na(ReADI_CBG_2015_Imp$B25088_002_T))
sum(is.na(ReADI_CBG_2015_Imp$B25088_002))

ReADI_CBG_2015_Imp$GEOID_CT <- NULL
ReADI_CBG_2015_Imp$B19013_001_T <- NULL
ReADI_CBG_2015_Imp$B25064_001_T <- NULL
ReADI_CBG_2015_Imp$B25077_001_T <- NULL
ReADI_CBG_2015_Imp$B25088_002_T <- NULL
sum(is.na(ReADI_CBG_2015_Imp)) # 0
save(ReADI_CBG_2015_Imp, file = "ReADI_CBG_2015_Imp.RData")

sum(is.na(ReADI_CBG_2015_Imp))
sum(is.na(ReADI_CT_2015_Imp))
sum(is.na(ReADI_C_2015_Imp))
```
  </div>
</div>

<br><br>

# ReADI Production

<br>

## Method

<br>

These steps are done separately at the county, census tract, and census block group levels. 

<strong><strong>1.</strong></strong> Using the imputed census data taken at each geography the 17 variables will be calculated. Below is a table summarizing the variables, census tables, and formulas we will be using to calculate the ReADI:

<div style="margin:16px 0">
 <center><button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code1111"><h4 class="text-white">CBG ReADI Variables</h4></button></center>
  <div id="code1111" class="collapse">
+-----------------------------------------------------------------------+---------+-------+:----------------------------------------------:+
|Census Variable                                                        |Name     |Table  |<center>Formula</center>                        |
+=======================================================================+=========+=======+:==============================================:+
|Median family income, \$                                               |MEDFI    |B19013 |$001$                                           |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Income disparity$^a$                                                   |INCDIS   |B19001 |$ln({\sum(002-004)+1\over\sum(014-017)+1}x100)$|
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Families below poverty level, %                                        |BELPL    |B17017 |${002\over001}x100$                             |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Population below 150% of the poverty threshold, %                      |PL150    |C17002 |${\sum(002-005)\over 001}x100$                  |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Population ≥25y with < a high school diploma, %                        |EDU9YR   |B15003 |${\sum(002-016)\over 001}x100$                  |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Population ≥25y with at least a bachelor's degree, %                   |HSDPNHIGH|B15003 |${\sum(022-025)\over 001}x100$                  |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Employed persons ≥16y in white-collar occupations, %                   |EMPWC    |C24010 |${\sum(003,027,039,063)\over 001}x100$          |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Civilian labor force population ≥16 y unemployed, %                    |UNEMP    |B23025 |${005\over003}x100$                             |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Median home value, \$                                                  |MEDHV    |B25077 |$001$                                           |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Median gross rent, \$                                                  |MEDRNT   |B25064 |$001$                                           |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Median monthly mortgage, \$                                            |MEDMORT  |B25088 |$002$                                           |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Owner-occupied housing units, %                                        |OWN      |B25003 |${002\over001}x100$                             |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Single-parent households with children <18y, %                         |SNGPNT   |B11005 |${\sum(006-007)\over 001}x100$                  |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Households without a motor vehicle, %                                  |NOVEH    |B25044 |${\sum(010,003)\over 001}x100$                  |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Households without internet, %                                         |NOINT    |B25043 |${\sum(007,016)\over 001}x100$                  |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Occupied housing units without complete plumbing, %                    |NOPLUM   |B25049 |$ln({\sum(004,007)+1\over001+1}x100)$          |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
|Households with more than 1 person per room, %                         |CROWD    |B25014 |${\sum(005-007,011-013)\over 001}x100$          |
+-----------------------------------------------------------------------+---------+-------+------------------------------------------------+
*<sup>$^a$ Income disparity in 2015 was defined as the log of 100×ratio of number of households with <\$20,000 income to number of households with ≥\$100,000 income.</sup>*
  </div>
</div>

<strong><strong>2.</strong></strong> Resulting NAs are converted to 0 as any found after calculation would be due to both the numerator and denominator being 0 and so will be converted to 0 as there are no true NAs in the data.

<strong><strong>3.</strong></strong> The 17 calculated variables are standardized via zscore.

<strong><strong>4.</strong></strong> Population weights are calculated using census variable B01003_001 (Universe: Total population) by dividing the areas population from the total population represented in the data labelled as POPWT. This value is then transformed where the square root + 1 is applied to account for the vastly different distributions across geographies and large skews to make them more comparable to one another.  

<strong><strong>5.</strong></strong> Factor analysis is applied (fa function from the stats package) where population weights are applied, a 1 factor solution requested, and the factor method used is the principal factor solution. All other options were default.

<br><br>

> CBG

<br>

<div class = "row">
<div class = "col-md-3">
|Variable  |NAs |Weights|
|:---------|:---|------:|
|MEDFI     |0   |-0.268 |
|INCDIS    |0   |0.166  |
|BELPL     |429 |0.077  |
|PL150     |362 |0.230  |
|EDU9YR    |37  |0.073  |
|HSDPNHIGH |37  |-0.106 |
|EMPWC     |265 |-0.107 |
|UNEMP     |249 |0.026  |
|MEDHV     |0   |-0.028 |
|MEDRNT    |0   |-0.037 |
|MEDMORT   |0   |-0.010 |
|OWN       |429 |0.013  |
|SNGPNT    |429 |0.059  |
|NOVEH     |429 |0.012  |
|NOINT     |429 |0.024  |
|NOPLUM    |0   |0.013  |
|CROWD     |429 |0.029  |
</div>

<div class = "col-md-1">
</div>

<div class = "col-md-6">
The table describes the amount of NAs which resulted from the 17 variable calculations and which were subsequently converted to 0 and the weights resulting from factor analysis. The code produced several products, the zscore standardized 17 composite variables, a raw ReADI score which can be used to subset down to whichever geographical area is relevant for a user's purposed and then re-scaled from 0-100, and then a nationally representative ADI scaled from 0-100. The ReADI_2015_CBG_RawVars.RData file has the 17 variables prior to zscore them.
</div>

<div class = "col-md-2">
</div>
</div>


<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code1019">Show Code</button>
  <div id="code1019" class="collapse">
```{r, eval = F}
load("ReADI_CBG_2015_Imp.RData")

ReADI_CBG_2015 <- dplyr::mutate(ReADI_CBG_2015_Imp,
    MEDFI = B19013_001,
    INCDIS = log(100*((B19001_002 + B19001_003 + B19001_004 + 1)/(B19001_014 + B19001_015 + B19001_016 + B19001_017 + 1))),
    BELPL = B17017_002/B17017_001*100,
    PL150 = (C17002_005 + C17002_004 + C17002_003 + C17002_002)/C17002_001*100,
    EDU9YR = (B15003_002 + B15003_003 + B15003_004 + B15003_005 + B15003_006 + B15003_007 + B15003_008 + B15003_009 + B15003_010 + B15003_011 + B15003_012 + B15003_013 + B15003_014 + B15003_015 + B15003_016)/B15003_001*100,
    HSDPNHIGH = (B15003_022 + B15003_023 + B15003_024 + B15003_025)/B15003_001*100,
    EMPWC = (C24010_003 + C24010_027 + C24010_039 + C24010_063)/ C24010_001*100,
    UNEMP = B23025_005/B23025_003*100,
    MEDHV = B25077_001,
    MEDRNT = B25064_001,
    MEDMORT = B25088_002,
    OWN = B25003_002/B25003_001*100,
    SNGPNT = (B11005_006 +  B11005_007)/B11005_001*100,
    NOVEH = (B25044_010 + B25044_003)/B25044_001*100,
    NOINT = (B25043_007 + B25043_016)/B25043_001*100,
    NOPLUM = log((B25049_007 + B25049_004 + 1)/(B25049_001 + 1)*100),
    CROWD = (B25014_005 + B25014_006 + B25014_007 + B25014_011 + B25014_012 + B25014_013)/B25014_001*100,
    POPWT = B01003_001
) %>%
  select(GEOID, Block_Group, Tract, County, State, MEDFI, INCDIS, BELPL, PL150, EDU9YR, HSDPNHIGH, EMPWC, UNEMP, MEDHV, MEDRNT, MEDMORT, OWN, SNGPNT, NOVEH, NOINT, NOPLUM, CROWD, POPWT
)
hist(ReADI_CBG_2015$POPWT)
ReADI_CBG_2015$POPWT <- log(ReADI_CBG_2015$POPWT)

# Replacing NA values resulting from 0 values back to 0. 
colSums(is.na(ReADI_CBG_2015))
ReADI_CBG_2015$BELPL <- ifelse(is.na(ReADI_CBG_2015$BELPL), 0, ReADI_CBG_2015$BELPL)
ReADI_CBG_2015$PL150 <- ifelse(is.na(ReADI_CBG_2015$PL150), 0, ReADI_CBG_2015$PL150)
ReADI_CBG_2015$EDU9YR <- ifelse(is.na(ReADI_CBG_2015$EDU9YR), 0, ReADI_CBG_2015$EDU9YR)
ReADI_CBG_2015$HSDPNHIGH <- ifelse(is.na(ReADI_CBG_2015$HSDPNHIGH), 0, ReADI_CBG_2015$HSDPNHIGH)
ReADI_CBG_2015$EMPWC <- ifelse(is.na(ReADI_CBG_2015$EMPWC), 0, ReADI_CBG_2015$EMPWC)
ReADI_CBG_2015$UNEMP <- ifelse(is.na(ReADI_CBG_2015$UNEMP), 0, ReADI_CBG_2015$UNEMP)
ReADI_CBG_2015$OWN <- ifelse(is.na(ReADI_CBG_2015$OWN), 0, ReADI_CBG_2015$OWN)
ReADI_CBG_2015$SNGPNT <- ifelse(is.na(ReADI_CBG_2015$SNGPNT), 0, ReADI_CBG_2015$SNGPNT)
ReADI_CBG_2015$NOVEH <- ifelse(is.na(ReADI_CBG_2015$NOVEH), 0, ReADI_CBG_2015$NOVEH)
ReADI_CBG_2015$NOINT <- ifelse(is.na(ReADI_CBG_2015$NOINT), 0, ReADI_CBG_2015$NOINT)
ReADI_CBG_2015$CROWD <- ifelse(is.na(ReADI_CBG_2015$CROWD), 0, ReADI_CBG_2015$CROWD)
ReADI_2015_CBG_RawVars <- ReADI_CBG_2015
save(ReADI_2015_CBG_RawVars, file = "ReADI_2015_CBG_RawVars.RData")

ADI_short <- ReADI_CBG_2015[, c("GEOID", "MEDFI", "INCDIS", "BELPL", "PL150", "EDU9YR", "HSDPNHIGH", "EMPWC", "UNEMP", "MEDHV", "MEDRNT", "MEDMORT", "OWN", "SNGPNT", "NOVEH", "NOINT", "NOPLUM", "CROWD")]
rownames(ADI_short) <- ADI_short$GEOID
ADI_short$GEOID <- NULL

# zscore transform
ADI_short <- as.data.frame(scale(ADI_short))

# Checking NAs and Inf values
NA_INF_CBG <- as.data.frame(t(sapply(1:ncol(ADI_short), function(x){
  NAs <- sum(is.na(ADI_short[,x]))
  Infs <- sum(is.infinite(ADI_short[,x]))
  c(NAs, Infs)
})))
colnames(NA_INF_CBG) <- c("NA", "Infinite")
rownames(NA_INF_CBG) <- colnames(ADI_short)

# Factor analysis is performed with a one factor solution which will be the scores applied to standardized data.
identical(rownames(ADI_short), ReADI_CBG_2015$GEOID)
adi.fa <- fa(ADI_short, nfactors = 1, fm = "pa", weight = ReADI_CBG_2015$POPWT)
ReADI_2015_Wt <- as.data.frame(adi.fa$weights)
colnames(ReADI_2015_Wt) <- "CBG"
ReADI_2015_Wt$Vars <- rownames(ReADI_2015_Wt)
ReADI_2015_Wt <- ReADI_2015_Wt[,c(2,1)]
save(ReADI_2015_Wt, file = "ReADI_2015_Wt.RData")

# Applying new weights coefficients to zscore data
identical(rownames(ReADI_2015_Wt), colnames(ADI_short))
ReADI_CBG_2015 <- as.data.frame(do.call(cbind,lapply(1:nrow(ReADI_2015_Wt), function(x){
  ADI_short[,x]*ReADI_2015_Wt[x,"CBG"]
})))
colnames(ReADI_CBG_2015) <- colnames(ADI_short)
rownames(ReADI_CBG_2015) <- rownames(ADI_short)

# Summing the weighted measures per CBG
ReADI_CBG_2015$ReADI_CBG_Raw <- rowSums(ReADI_CBG_2015, na.rm = F)
sum(is.na(ReADI_CBG_2015$ReADI_CBG_Raw)) # 0
ReADI_CBG_2015$GEOID <- rownames(ReADI_CBG_2015)

# There is subsequent percentile ranking done for relative comparisons of CBG nationally.
ReADI_CBG_2015$ReADI_CBG_NR <- round(percent_rank(ReADI_CBG_2015$ReADI_CBG_Raw)*100, 0)

# Range of variables after weights applied
Range_CBG <- as.data.frame(t(sapply(1:17, function(x){
  Mid <- median(ReADI_CBG_2015[,x], na.rm = T)
  Min <- range(ReADI_CBG_2015[,x], na.rm = T)[1]
  Max <- range(ReADI_CBG_2015[,x], na.rm = T)[2]
  Dif <- Max - Min
  c(Mid, Min, Max, Dif)
})))
colnames(Range_CBG) <- c("Median", "Min", "Max", "Dif")
rownames(Range_CBG) <- colnames(ReADI_CBG_2015)[1:17]
save(ReADI_CBG_2015, file = "ReADI_CBG_2015.RData")
```
  </div>
</div>

<br><br>

> CT

<br>

<div class = "row">
<div class = "col-md-3">
|Variable  |NAs |Weights|
|:---------|:---|------:|
|MEDFI     |0   |-0.297 |
|INCDIS    |0   |0.242  |
|BELPL     |175 |0.011  |
|PL150     |143 |0.286  |
|EDU9YR    |10  |0.062  |
|HSDPNHIGH |10  |-0.041 |
|EMPWC     |114 |-0.136 |
|UNEMP     |110 |0.012  |
|MEDHV     |0   |-0.038 |
|MEDRNT    |0   |0.002  |
|MEDMORT   |0   |0.060  |
|OWN       |175 |0.063  |
|SNGPNT    |175 |0.083  |
|NOVEH     |175 |-0.003 |
|NOINT     |175 |0.026  |
|NOPLUM    |0   |0.011  |
|CROWD     |175 |0.015  |
</div>

<div class = "col-md-1">
</div>

<div class = "col-md-6">
The table describes the amount of NAs which resulted from the 17 variable calculations and which were subsequently converted to 0 and the weights resulting from factor analysis. The code produced several products, the zscore standardized 17 composite variables, a raw ReADI score which can be used to subset down to whichever geographical area is relevant for a user's purposed and then re-scaled from 0-100, and then a nationally representative ADI scaled from 0-100. The ReADI_2015_CT_RawVars.RData file has the 17 variables prior to zscore them.
</div>

<div class = "col-md-2">
</div>
</div>


<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code1340">Show Code</button>
  <div id="code1340" class="collapse">
```{r, eval = F}
load("ReADI_CT_2015_Imp.RData")

ReADI_CT_2015 <- dplyr::mutate(ReADI_CT_2015_Imp,
    MEDFI = B19013_001,
    INCDIS = log(100*((B19001_002 + B19001_003 + B19001_004 + 1)/(B19001_014 + B19001_015 + B19001_016 + B19001_017 + 1))),
    BELPL = B17017_002/B17017_001*100,
    PL150 = (C17002_005 + C17002_004 + C17002_003 + C17002_002)/C17002_001*100,
    EDU9YR = (B15003_002 + B15003_003 + B15003_004 + B15003_005 + B15003_006 + B15003_007 + B15003_008 + B15003_009 + B15003_010 + B15003_011 + B15003_012 + B15003_013 + B15003_014 + B15003_015 + B15003_016)/B15003_001*100,
    HSDPNHIGH = (B15003_022 + B15003_023 + B15003_024 + B15003_025)/B15003_001*100,
    EMPWC = (C24010_003 + C24010_027 + C24010_039 + C24010_063)/ C24010_001*100,
    UNEMP = B23025_005/B23025_003*100,
    MEDHV = B25077_001,
    MEDRNT = B25064_001,
    MEDMORT = B25088_002,
    OWN = B25003_002/B25003_001*100,
    SNGPNT = (B11005_006 +  B11005_007)/B11005_001*100,
    NOVEH = (B25044_010 + B25044_003)/B25044_001*100,
    NOINT = (B25043_007 + B25043_016)/B25043_001*100,
    NOPLUM = log((B25049_007 + B25049_004 + 1)/(B25049_001 + 1)*100),
    CROWD = (B25014_005 + B25014_006 + B25014_007 + B25014_011 + B25014_012 + B25014_013)/B25014_001*100,
    POPWT = B01003_001
) %>%
  select(GEOID, Tract, County, State, MEDFI, INCDIS, BELPL, PL150, EDU9YR, HSDPNHIGH, EMPWC, UNEMP, MEDHV, MEDRNT, MEDMORT, OWN, SNGPNT, NOVEH, NOINT, NOPLUM, CROWD, POPWT
)
hist(ReADI_CT_2015$POPWT)
ReADI_CT_2015$POPWT <- log(ReADI_CT_2015$POPWT)

# Replacing NA values resulting from 0 values back to 0. 
colSums(is.na(ReADI_CT_2015))
ReADI_CT_2015$BELPL <- ifelse(is.na(ReADI_CT_2015$BELPL), 0, ReADI_CT_2015$BELPL)
ReADI_CT_2015$PL150 <- ifelse(is.na(ReADI_CT_2015$PL150), 0, ReADI_CT_2015$PL150)
ReADI_CT_2015$EDU9YR <- ifelse(is.na(ReADI_CT_2015$EDU9YR), 0, ReADI_CT_2015$EDU9YR)
ReADI_CT_2015$HSDPNHIGH <- ifelse(is.na(ReADI_CT_2015$HSDPNHIGH), 0, ReADI_CT_2015$HSDPNHIGH)
ReADI_CT_2015$EMPWC <- ifelse(is.na(ReADI_CT_2015$EMPWC), 0, ReADI_CT_2015$EMPWC)
ReADI_CT_2015$UNEMP <- ifelse(is.na(ReADI_CT_2015$UNEMP), 0, ReADI_CT_2015$UNEMP)
ReADI_CT_2015$OWN <- ifelse(is.na(ReADI_CT_2015$OWN), 0, ReADI_CT_2015$OWN)
ReADI_CT_2015$SNGPNT <- ifelse(is.na(ReADI_CT_2015$SNGPNT), 0, ReADI_CT_2015$SNGPNT)
ReADI_CT_2015$NOVEH <- ifelse(is.na(ReADI_CT_2015$NOVEH), 0, ReADI_CT_2015$NOVEH)
ReADI_CT_2015$NOINT <- ifelse(is.na(ReADI_CT_2015$NOINT), 0, ReADI_CT_2015$NOINT)
ReADI_CT_2015$CROWD <- ifelse(is.na(ReADI_CT_2015$CROWD), 0, ReADI_CT_2015$CROWD)
ReADI_2015_CT_RawVars <- ReADI_CT_2015
save(ReADI_2015_CT_RawVars, file = "ReADI_2015_CT_RawVars.RData")

ADI_short <- ReADI_CT_2015[, c("GEOID", "MEDFI", "INCDIS", "BELPL", "PL150", "EDU9YR", "HSDPNHIGH", "EMPWC", "UNEMP", "MEDHV", "MEDRNT", "MEDMORT", "OWN", "SNGPNT", "NOVEH", "NOINT", "NOPLUM", "CROWD")]
rownames(ADI_short) <- ADI_short$GEOID
ADI_short$GEOID <- NULL

# zscore transform
ADI_short <- as.data.frame(scale(ADI_short))

# Checking NAs and Inf values
NA_INF_CT <- as.data.frame(t(sapply(1:ncol(ADI_short), function(x){
  NAs <- sum(is.na(ADI_short[,x]))
  Infs <- sum(is.infinite(ADI_short[,x]))
  c(NAs, Infs)
})))
colnames(NA_INF_CT) <- c("NA", "Infinite")
rownames(NA_INF_CT) <- colnames(ADI_short)

# Factor analysis is performed with a one factor solution which will be the scores applied to standardized data.
load("ReADI_2015_Wt.RData")
identical(rownames(ADI_short), ReADI_CT_2015$GEOID)
adi.fa <- fa(ADI_short, nfactors = 1, fm = "pa", weight = ReADI_CT_2015$POPWT)
identical(colnames(adi.fa$residual),ReADI_2015_Wt$Vars)
ReADI_2015_Wt$CT <- adi.fa$weights[,"PA1"]
save(ReADI_2015_Wt, file = "ReADI_2015_Wt.RData")

# Applying new weights coefficients to zscore data
identical(rownames(ReADI_2015_Wt), colnames(ADI_short))
ReADI_CT_2015 <- as.data.frame(do.call(cbind,lapply(1:nrow(ReADI_2015_Wt), function(x){
  ADI_short[,x]*ReADI_2015_Wt[x,"CT"]
})))
colnames(ReADI_CT_2015) <- colnames(ADI_short)
rownames(ReADI_CT_2015) <- rownames(ADI_short)

# Summing the weighted measures per CBG
ReADI_CT_2015$ReADI_CT_Raw <- rowSums(ReADI_CT_2015, na.rm = F)
sum(is.na(ReADI_CT_2015$ReADI_CT_Raw)) # 0
ReADI_CT_2015$GEOID <- rownames(ReADI_CT_2015)

# There is subsequent percentile ranking done for relative comparisons of CBG nationally.
ReADI_CT_2015$ReADI_CT_NR <- round(percent_rank(ReADI_CT_2015$ReADI_CT_Raw)*100, 0)

# Range of variables after weights applied
Range_CT <- as.data.frame(t(sapply(1:17, function(x){
  Mid <- median(ReADI_CT_2015[,x], na.rm = T)
  Min <- range(ReADI_CT_2015[,x], na.rm = T)[1]
  Max <- range(ReADI_CT_2015[,x], na.rm = T)[2]
  Dif <- Max - Min
  c(Mid, Min, Max, Dif)
})))
colnames(Range_CT) <- c("Median", "Min", "Max", "Dif")
rownames(Range_CT) <- colnames(ReADI_CT_2015)[1:17]
save(ReADI_CT_2015, file = "ReADI_CT_2015.RData")
```
  </div>
</div>

<br><br>

> C

<br>

<div class = "row">
<div class = "col-md-3">
|Variable  |NAs |Weights|
|:---------|:---|------:|
|MEDFI     |0   |-0.233 |
|INCDIS    |0   |0.310  |
|BELPL     |0   |0.028  |
|PL150     |0   |0.194  |
|EDU9YR    |0   |0.041  |
|HSDPNHIGH |0   |-0.121 |
|EMPWC     |0   |-0.074 |
|UNEMP     |0   |-0.029 |
|MEDHV     |0   |-0.096 |
|MEDRNT    |0   |-0.018 |
|MEDMORT   |0   |0.022  |
|OWN       |0   |0.039  |
|SNGPNT    |0   |0.079  |
|NOVEH     |0   |-0.029 |
|NOINT     |0   |0.024  |
|NOPLUM    |0   |0.013  |
|CROWD     |0   |0.051  |
</div>

<div class = "col-md-1">
</div>

<div class = "col-md-6">
The table describes the amount of NAs which resulted from the 17 variable calculations and which were subsequently converted to 0 and the weights resulting from factor analysis. The code produced several products, the zscore standardized 17 composite variables, a raw ReADI score which can be used to subset down to whichever geographical area is relevant for a user's purposed and then re-scaled from 0-100, and then a nationally representative ADI scaled from 0-100. The ReADI_2015_C_RawVars.RData file has the 17 variables prior to zscore them.
</div>

<div class = "col-md-2">
</div>
</div>

<div style="margin:16px 0">
  <button type="button" class="btn btn-dark" control-id="ControlID-54" data-bs-toggle="collapse" data-bs-target="#code1488">Show Code</button>
  <div id="code1488" class="collapse">
```{r, eval = F}
load("ReADI_C_2015_Imp.RData")

ReADI_C_2015 <- dplyr::mutate(ReADI_C_2015_Imp,
    MEDFI = B19013_001,
    INCDIS = log(100*((B19001_002 + B19001_003 + B19001_004 + 1)/(B19001_014 + B19001_015 + B19001_016 + B19001_017 + 1))),
    BELPL = B17017_002/B17017_001*100,
    PL150 = (C17002_005 + C17002_004 + C17002_003 + C17002_002)/C17002_001*100,
    EDU9YR = (B15003_002 + B15003_003 + B15003_004 + B15003_005 + B15003_006 + B15003_007 + B15003_008 + B15003_009 + B15003_010 + B15003_011 + B15003_012 + B15003_013 + B15003_014 + B15003_015 + B15003_016)/B15003_001*100,
    HSDPNHIGH = (B15003_022 + B15003_023 + B15003_024 + B15003_025)/B15003_001*100,
    EMPWC = (C24010_003 + C24010_027 + C24010_039 + C24010_063)/ C24010_001*100,
    UNEMP = B23025_005/B23025_003*100,
    MEDHV = B25077_001,
    MEDRNT = B25064_001,
    MEDMORT = B25088_002,
    OWN = B25003_002/B25003_001*100,
    SNGPNT = (B11005_006 +  B11005_007)/B11005_001*100,
    NOVEH = (B25044_010 + B25044_003)/B25044_001*100,
    NOINT = (B25043_007 + B25043_016)/B25043_001*100,
    NOPLUM = log((B25049_007 + B25049_004 + 1)/(B25049_001 + 1)*100),
    CROWD = (B25014_005 + B25014_006 + B25014_007 + B25014_011 + B25014_012 + B25014_013)/B25014_001*100,
    POPWT = B01003_001
) %>%
  select(GEOID, County, State, MEDFI, INCDIS, BELPL, PL150, EDU9YR, HSDPNHIGH, EMPWC, UNEMP, MEDHV, MEDRNT, MEDMORT, OWN, SNGPNT, NOVEH, NOINT, NOPLUM, CROWD, POPWT, 
) 
hist(ReADI_C_2015$POPWT)
ReADI_C_2015$POPWT <- log(ReADI_C_2015$POPWT)


# Replacing NA values resulting from 0 values back to 0. 
colSums(is.na(ReADI_C_2015))
ReADI_2015_C_RawVars <- ReADI_C_2015
save(ReADI_2015_C_RawVars, file = "ReADI_2015_C_RawVars.RData")

ADI_short <- ReADI_C_2015[, c("GEOID", "MEDFI", "INCDIS", "BELPL", "PL150", "EDU9YR", "HSDPNHIGH", "EMPWC", "UNEMP", "MEDHV", "MEDRNT", "MEDMORT", "OWN", "SNGPNT", "NOVEH", "NOINT", "NOPLUM", "CROWD")]
rownames(ADI_short) <- ADI_short$GEOID
ADI_short$GEOID <- NULL

# zscore transform
ADI_short <- as.data.frame(scale(ADI_short))

# Checking NAs and Inf values
NA_INF_C <- as.data.frame(t(sapply(1:ncol(ADI_short), function(x){
  NAs <- sum(is.na(ADI_short[,x]))
  Infs <- sum(is.infinite(ADI_short[,x]))
  c(NAs, Infs)
})))
colnames(NA_INF_C) <- c("NA", "Infinite")
rownames(NA_INF_C) <- colnames(ADI_short)

# Factor analysis is performed with a one factor solution which will be the scores applied to standardized data.
load("ReADI_2015_Wt.RData")
identical(rownames(ADI_short), ReADI_C_2015$GEOID)
adi.fa <- fa(ADI_short, nfactors = 1, fm = "pa", weight = ReADI_C_2015$POPWT)
identical(colnames(adi.fa$residual),ReADI_2015_Wt$Vars)
ReADI_2015_Wt$C <- adi.fa$weights[,"PA1"]
save(ReADI_2015_Wt, file = "ReADI_2015_Wt.RData")
write.csv(ReADI_2015_Wt, file = "ReADI_2015_Wt.csv")

# Applying new weights coefficients to zscore data
identical(rownames(ReADI_2015_Wt), colnames(ADI_short))
ReADI_C_2015 <- as.data.frame(do.call(cbind,lapply(1:nrow(ReADI_2015_Wt), function(x){
  ADI_short[,x]*ReADI_2015_Wt[x,"C"]
})))
colnames(ReADI_C_2015) <- colnames(ADI_short)
rownames(ReADI_C_2015) <- rownames(ADI_short)

# Summing the weighted measures per CBG
ReADI_C_2015$ReADI_C_Raw <- rowSums(ReADI_C_2015, na.rm = F)
sum(is.na(ReADI_C_2015$ReADI_C_Raw)) # 0
ReADI_C_2015$GEOID <- rownames(ReADI_C_2015)

# There is subsequent percentile ranking done for relative comparisons of CBG nationally.
ReADI_C_2015$ReADI_C_NR <- round(percent_rank(ReADI_C_2015$ReADI_C_Raw)*100, 0)

# Range of variables after weights applied
Range_C <- as.data.frame(t(sapply(1:17, function(x){
  Mid <- median(ReADI_C_2015[,x], na.rm = T)
  Min <- range(ReADI_C_2015[,x], na.rm = T)[1]
  Max <- range(ReADI_C_2015[,x], na.rm = T)[2]
  Dif <- Max - Min
  c(Mid, Min, Max, Dif)
})))
colnames(Range_C) <- c("Median", "Min", "Max", "Dif")
rownames(Range_C) <- colnames(ReADI_C_2015)[1:17]
save(ReADI_C_2015, file = "ReADI_C_2015.RData")
```
  </div>
</div>

<br><br>
